Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS

### 2011-06-08 ms.tex
Minor changes, new skymaps and sections

The Coma cluster of galaxies (ACO 1656) is one of the most thouroughly studied clusters over all wavelengths \citep{article:Voges_etal:1999}. Located at a distance of about 100 Mpc \citep[$z=0.023$;][]{article:StrubleRood:1999}, it is one of the closest and also one of the most massive clusters \citep[$M \sim 10^{15}M_{\odot}$;][]{article:Smith:1983, article:Kubo_etal:2008}. It hosts both a giant radio halo \citep{article:Giovannini_etal:1993,article:Thierbach_etal:2003} and peripheral radio relic, which appears connected to the radio halo with a ``diffuse'' bridge \citep[see discussion in][]{article:BrownRudnick:2010}. It has been suggested that the relic is an infall shock. Extended soft X-ray (SXR) emission is evident from the ROSAT All-sky survey in the 0.1 to 2.4 keV band \citep{article:BrielHenryBohringer:1992} and observations with XMM-Newton \citep{article:Briel_etal:2001} revealed substructure in the X-ray halo. As such it is a natural candidate for gamma-ray observations. In this article, results from the VERITAS observations of the Coma cluster of galaxies are reported, together with its constraints on cosmic-ray particle populations, magnetic fields, and dark matter in the cluster.

@@

The Coma cluster was observed with VERITAS between March and May in 2008 with all four telescopes fully operational. The total exposure amounts to 18.6 hours of quality-selected live time, i.e. time periods of astronomical darkness, clear sky conditions and without technical problems with the array. The center of the Coma cluster was tracked in \emph{wobble} mode, where the expected source location is offset from the center of the field of view, to allow for simultaneous background estimation \citep{article:Formin_etal:1994}. All of the observations were made in a small range of zenith angles averaging $\sim 21^{\circ}$.

@@

Gamma-ray like events are separated from the CR background by imposing cuts on scaled parameters \citep{article:Aharonian_etal:1997, article:Krawczynski_etal:2006} calculated from a parameterized moment analysis of the shower images \citep{inproc:Hillas:1985}. The gamma-hadron separation cuts used in this analysis were optimized a priori for a weak point source (3\% Crab Nebula flux level) and a photon index of -2.4 using data taken on the Crab Nebula during the same epoch and are termed \emph{standard} cuts. Because the VHE gamma-ray spectrum for the Coma cluster is expected to be hard, photon index of about -2.3, these standard cuts are suitable.

The Coma cluster is a complex region with many plausible sites for gamma-ray emission.

The ring background model \citep{article:Aharonian_etal:2001} was used to produce skymaps of the VERITAS excess over the field of view. The total number of events in a given region of interest is then compared to the estimated background from the off-source region scaled by the ratio of the solid angles to produce a final excess.

%The observed radio halo and thermal X-ray emission in the Coma cluster suggest that the VHE gamma-ray emission should have a similar spatial extension.

\begin{figure*}
\begin{center}
\scalebox{0.5}{\plotone{f1a.eps}}
\scalebox{0.5}{\plotone{f1b.eps}}
\end{center}
\caption{Skymaps.}
\label{fig:skymaps}
\end{figure*}

\section{Results and Discussion}
No significant excess was detected with VERITAS in the Coma cluster.
A 99\% confidence level upper limit is calculated for each region of interest using the method described by \citet{article:Rolke:2005} and assuming Gaussian background.

@@

>>>>

The Coma cluster of galaxies (ACO 1656) is one of the most thouroughly studied clusters over all wavelengths \citep{article:Voges_etal:1999}. Located at a distance of about 100 Mpc \citep[$z=0.023$;][]{article:StrubleRood:1999}, it is one of the closest and also one of the most massive clusters \citep[$M \sim 10^{15}M_{\odot}$;][]{article:Smith:1983, article:Kubo_etal:2008}. It hosts both a giant radio halo \citep{article:Giovannini_etal:1993,article:Thierbach_etal:2003} and peripheral radio relic, which appears connected to the radio halo with a ``diffuse'' bridge \citep[see discussion in][]{article:BrownRudnick:2010}. It has been suggested that the relic is an infall shock. Extended soft X-ray (SXR) emission is evident from the ROSAT all-sky survey in the 0.1 to 2.4 keV band \citep{article:BrielHenryBohringer:1992} and observations with XMM-Newton \citep{article:Briel_etal:2001} revealed substructure in the X-ray halo. As such it is a natural candidate for gamma-ray observations. In this article, results from the VERITAS observations of the Coma cluster of galaxies are reported, together with its constraints on cosmic-ray particle populations, magnetic fields, and dark matter in the cluster.

@@

The Coma cluster was observed with VERITAS between March and May in 2008 with all four telescopes fully operational. The total exposure amounts to 18.6 hours of quality-selected live time, i.e. time periods of astronomical darkness, clear sky conditions and without technical problems with the array. The center of the cluster was tracked in \emph{wobble} mode, where the expected source location is offset from the center of the field of view, to allow for simultaneous background estimation \citep{article:Formin_etal:1994}. All of the observations were made in a small range of zenith angles averaging $\sim 21^{\circ}$.

@@

Gamma-ray like events are separated from the CR background by imposing cuts on scaled parameters \citep{article:Aharonian_etal:1997, article:Krawczynski_etal:2006} calculated from a parameterized moment analysis of the shower images \citep{inproc:Hillas:1985}. The gamma-hadron separation cuts used in this analysis were optimized a priori for a weak point source (3\% Crab Nebula flux level) and a photon index of -2.4 using data taken on the Crab Nebula during the same epoch and are termed \emph{standard} cuts. Because the VHE gamma-ray spectrum for the Coma cluster is expected to be hard, photon index of about -2.3, these standard cuts are suitable for the analysis of the Coma cluster data set.

The ring background model \citep{article:Aharonian_etal:2001} is used to estimate the background due to CRs. The total number of events in a given region of interest is then compared to the estimated background from the off-source region scaled by the ratio of the solid angles to produce a final excess. The VHE gamma-ray significance is then calculated using the method by \citet{article:LiMa:1983}.  Significance skymaps over the VERITAS field of view are shown in Figure \ref{fig:Coma:skymaps} with overlaid X-ray and radio contours from the ROSAT all-sky survey \citep{article:BrielHenryBohringer:1992}  and GBT 1.4 GHz observations \citep{article:BrownRudnick:2010} respectively.

\begin{figure*}
\begin{center}
\scalebox{0.5}{\plotone{f1a.eps}}
\scalebox{0.5}{\plotone{f1b.eps}}
\end{center}
\caption{Skymaps.}
\label{fig:Coma:skymaps}
\end{figure*}

\section{Results}
No significant excess was detected with VERITAS in the Coma cluster.
A 99\% confidence level upper limit is calculated for each region of interest using the method described by \citet{article:Rolke:2005} and assuming Gaussian background.

@@

\section{Cosmic Rays}

\section{Dark Matter}
